Dynamic Adaptive Saturated Sliding Mode Control For Deployment of Tethered Satellite System

### 2017-01-01 manuscript.tex
20170101

Figure \ref{fig:1} shows all the elements of the TSS via the geometric representation. In the figure, the axes of the coordinate frame $OXYZ$ passes through the center of the earth $O$. $OX$ and $OY$ are mutually perpendicular axes on the equatorial plane with $OX$ pointing to the spring equinox, and $OZ$ is perpendicular to the equatorial plane, bundling the axis of the earth upwards. The axis $OY$ completes the right-hand triad. The orbital coordinate frame is denoted as $O_2xyz$ with the origin $O_2$ located at the mass center of the TSS. $O_2x$ is directed in the direction of orbital velocity, meanwhile $O_2z$ is defined to point towards the center of the earth, generating the axis $O_2y$ by utilizing the right-hand triad. The origin of the TSS coordinate frame $O_2x'y'z'$ is also fixed at the mass center of the TSS, and the axis $O_2z'$ is directed along the tether and points to the subsatellite. The orientation of $O_2xyz$ relative to the coordinate frame $O_2x'y'z'$ indicates the in-plane and out-of-plane dynamics, where $\theta$ and $\phi$ stand for the in-plane angle and out-of-plane angle, respectively. $\bm R_0$ denotes the position vector between $O_2$ and $O$.

@@

and it's notable that the function above is saturated and owns an important property: $\vert \rm{Sine}\it{(x)}\vert\le\vert x\vert$, $\forall x$. The geometric illustration is shown in Figure~\ref{fig:2}, from which an interesting property can be found by sliding the curve $y=\rm{Sine}(\it{x})$ along the axis $x=0$ in the range $[-1,1]$, and once the function $y=\rm{Sine}\it{(x)}+\Delta h$ with $\vert\Delta h\vert<1$ settles, one can choose the function $y = x+\arcsin(\Delta h)$ to obtain following relations:

@@

Based on the expression above, one can obtain $\bar{m}\Omega^2L\approx 4.79\times 10^{-3}$, and suitable boundary values $\hat\tau_{tmin}$ and $\hat\tau_{tmax}$ can be chosen to make the dimensionless tension $\hat\tau_t$ fall in the region $[-3,0.01]$.\par
At this point, the simulations of the deployment dynamics of the TSS can be carried on. We firstly conduct the simulation about the pure deployment dynamics without any uncertainties to verify the effectiveness of the proposed control laws, and the results are illustrated in Figure~\ref{fig:3} to~\ref{fig:6}. Figure~\ref{fig:3} shows the dynamics of the length in different control laws, and it is obvious that all the methods can realize the deployment in aspect of the tether length. ST-SMC and QST-SMC complete the deployment at about 4 $rad$. As shown in Figure~\ref{fig:4}, S-SMC and QST-SMC have bigger length rate than ST-SMC before 1.2 $rad$, indicating the curves of the tether length climb more rapidly at the beginning of the deployment. However, the length rate of ST-SMC doesn't decline dramatically like the other two methods after 1.2 $rad$. This slowly varying length rate yields a smooth trajectory of the deployment shown in Figure~\ref{fig:3}, and meanwhile the length rate converges to the origin earlier than S-SMC and QST-SMC.\par
The in-plane and out-of-plane angles reflect the swing dynamics of the tether during the deployment, the details of which are drawn in Figue~\ref{fig:5}. For ST-SMC, the in-plane dynamic motion is regulated to vertical status at 4 $rad$, and the in-plane angle converges to the origin in the subsequent time. Hence the coalition of the length and the swing dynamics imply that the whole deployment mission is completed at about 4 $rad$ by utilizing ST-SMC. S-SMC makes the tether swing to the desired position at a low convergent speed near the vertical direction during 2 $rad$ to 4 $rad$.\par
The inputs in S-SMC for dynamic model~(\ref{eq:dynamic model}) are exhibited in Figure~\ref{fig:6}, in which it is obvious that the limited situations occur frequently. Thanks to the scalable dimensionless transformation determined conservatively, the tension exerted on the tether falls in the reasonable region.\par

@@

\bm\tau_d&=\left(0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000)\right)^T,

@@

where $\hat\bullet$ stands for the measurement of the term $\bullet$ with uncertainties. $\varpi$ is the uncertainty boundary parameter, and here we select $\varpi=2$. $\bm\tau_d$ is the external disturbance meaning  space perturbations, such as gravity gradient effect, aerodynamic interference and solar radiation pressure torque.  In this simulation, the measured terms can be specified as: $\hat{\bar m}=1.1\bar m$ may mean some space debris affixed to the subsatellite; $\bar L=1.02L$ and $\bar \Omega=0.98\Omega$ mean some measurement errors happening to the total length and the orbital angular velocity. In Fig~\ref{fig:6}, although the dynamics have already been  stable, the tension is switching in the high frequency. It is commonplace in the sliding mode control law design, called chattering, since the signum function is used in the input, which may cause unmodeled high frequency dynamics, and for reducing the chattering, we introduce the boundary layer technique into the proposed control laws with the thickness of the boundary layer specified as a positive scalar $\nu=1\times 10^{-4}$ in this simulation~\cite{1039802}.

The simulation results of the deployment dynamics subjected to uncertainties are exhibited from Fig~\ref{fig:7} to Fig~\ref{fig:10}. In Fig~\ref{fig:7}, although there exist parametric uncertainties compared with the dynamics in Figs~\ref{fig:3}-\ref{fig:6}, QST-SMC still performs the deployment with the shortest rise time and the minimum overshoot among the three methods, and the tether is deployed to the desired position at about 8 $rad$ by observing the dynamics of the in-plane and out-of-plane angles from Fig~\ref{fig:9}. In addition, the result in Fig~\ref{fig:10} shows that the tension falls in the reasonable region, and meanwhile the chattering of the input in the stable status is reduced by utilizing the boundary layer technique. \textcolor{red}{Fig.~\ref{fig:11} shows the in-plane and out-of-plane thruster forces. From the results, we can find that all the forces are reasonable and limited in the boundary, and have less chattering by using the boundary layer technique. The thruster forces have the similar behaviors like the tension, and the chattering is almost eliminated in the final stage. Hence we can say that the boundary layer technique can indeed improve the inputs and repress the chattering.} The control precisions in Figs~\ref{fig:7}-\ref{fig:9} are almost same to Figs~\ref{fig:3}-\ref{fig:6}, which means the loss of precision caused by the boundary layer technique is limited. Moreover, the simulation results show that the good performance of proposed controller and the suitable boundary layer thickness guarantee that the system state error converges to the small vicinity of origin, which is consistent with the analyses in Theorem~\ref{thm:boundary}. Hence, the robustness of the proposed control laws can guarantee the stability of the deployment subjected to parametric uncertainties and external disturbances.
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig7_20161025.eps}
\caption{Length dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:7}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig8_20161025.eps}
\caption{Length rate dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:8}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig9_20161025.eps}
\caption{Swing dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:9}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_fig10_20161025.eps}
\caption{Tension produced by utilizing boundary layer technique}
\label{fig:10}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{paper4_response_20161230.eps}
\caption{Thruster force}
\label{fig:11}
\end{figure}
\section{Conclusion}\label{sec:5}
In this paper, the proposed adaptive sliding mode control laws provide the deployment dynamics of the TSS with a simplified solution for the reaching phase, which ignores the nonlinear nature of the model, and the new dimensionless transformation subjected to the limited input guarantees the effectiveness of the auxiliary Lyapunov function with adaptive terms. The simulation results show that the proposed methods adjust the length of the tether and the swing dynamics as desired, although there exists the input limitation during the deployment. Also, the robustness to parametric uncertainties and external disturbances is verified reasonably. According to the results mentioned above, combining error dynamics design approach, adaptive approach and nonlinear sliding mode surface design technology may provide a further insight into regulation of the deployment dynamics of the TSS.
\section{Acknowledgment}
This work is partially supported by the National Natural Science Foundation of China (No. 61104112, 61673009, 61503101, 91438202).
\section{References}
\bibliography{paper4_ref}
\bibliographystyle{elsarticle-num}
\end{document}

>>>>

Fig~\ref{fig:1} shows all the elements of the TSS via the geometric representation. In the figure, the axes of the coordinate frame $OXYZ$ passes through the center of the earth $O$. $OX$ and $OY$ are mutually perpendicular axes on the equatorial plane with $OX$ pointing to the spring equinox, and $OZ$ is perpendicular to the equatorial plane, bundling the axis of the earth upwards. The axis $OY$ completes the right-hand triad. The orbital coordinate frame is denoted as $O_2xyz$ with the origin $O_2$ located at the mass center of the TSS. $O_2x$ is directed in the direction of orbital velocity, meanwhile $O_2z$ is defined to point towards the center of the earth, generating the axis $O_2y$ by utilizing the right-hand triad. The origin of the TSS coordinate frame $O_2x'y'z'$ is also fixed at the mass center of the TSS, and the axis $O_2z'$ is directed along the tether and points to the subsatellite. The orientation of $O_2xyz$ relative to the coordinate frame $O_2x'y'z'$ indicates the in-plane and out-of-plane dynamics, where $\theta$ and $\phi$ stand for the in-plane angle and out-of-plane angle, respectively. $\bm R_0$ denotes the position vector between $O_2$ and $O$.

@@

and it's notable that the function above is saturated and owns an important property: $\vert \rm{Sine}\it{(x)}\vert\le\vert x\vert$, $\forall x$. The geometric illustration is shown in Fig~\ref{fig:2}, from which an interesting property can be found by sliding the curve $y=\rm{Sine}(\it{x})$ along the axis $x=0$ in the range $[-1,1]$, and once the function $y=\rm{Sine}\it{(x)}+\Delta h$ with $\vert\Delta h\vert<1$ settles, one can choose the function $y = x+\arcsin(\Delta h)$ to obtain following relations:

@@

Based on the expression above, one can obtain $\bar{m}\Omega^2L\approx 0.0573N$, and suitable boundary values $\hat\tau_{tmin}$ and $\hat\tau_{tmax}$ can be chosen to make the dimensionless tension $\hat\tau_t$ fall in the region $[-3,0.01]$.\par
At this point, the simulations of the deployment dynamics of the TSS can be carried on. We firstly conduct the simulation about the pure deployment dynamics without any uncertainties to verify the effectiveness of the proposed control laws, and the results are illustrated in Fig~\ref{fig:3} to~\ref{fig:6}. Fig~\ref{fig:3} shows the dynamics of the length in different control laws, and it is obvious that all the methods can realize the deployment in aspect of the tether length. ST-SMC and QST-SMC complete the deployment at about 4 $rad$. As shown in Fig~\ref{fig:4}, S-SMC and QST-SMC have bigger length rate than ST-SMC before 1.2 $rad$, indicating the curves of the tether length climb more rapidly at the beginning of the deployment. However, the length rate of ST-SMC doesn't decline dramatically like the other two methods after 1.2 $rad$. This slowly varying length rate yields a smooth trajectory of the deployment shown in Fig~\ref{fig:3}, and meanwhile the length rate converges to the origin earlier than S-SMC and QST-SMC.\par
The in-plane and out-of-plane angles reflect the swing dynamics of the tether during the deployment, the details of which are drawn in Fig~\ref{fig:5}. For ST-SMC, the in-plane dynamic motion is regulated to vertical status at 4 $rad$, and the in-plane angle converges to the origin in the subsequent time. Hence the coalition of the length and the swing dynamics imply that the whole deployment mission is completed at about 4 $rad$ by utilizing ST-SMC. S-SMC makes the tether swing to the desired position at a low convergent speed near the vertical direction during 2 $rad$ to 4 $rad$.\par
The inputs in S-SMC for dynamic model~(\ref{eq:dynamic model}) are exhibited in Fig~\ref{fig:6}, in which it is obvious that the limited situations occur frequently. Thanks to the scalable dimensionless transformation determined conservatively, the tension exerted on the tether falls in the reasonable region. In addition, combining with the results in Figs~\ref{fig:3}-\ref{fig:6},\ref{fig:11}, the system sates converge to desired values as the inputs are frequently switching, implying that the dynamic stability can be guaranteed by the proposed approach.\par

@@

\bm\tau_d&=\left[5\times 10^{-2}\sin(\it{v}),\rm 5\times \rm 10^{-2}\sin(\it{v}), \rm 5\times 10^{-2}\sin(\it{v})\right]^T,

@@

where $\hat\bullet$ stands for the measurement of the term $\bullet$ with uncertainties. $\varpi$ is the uncertainty boundary parameter, and here we select $\varpi=2$. $\bm\tau_d$ is the external disturbance meaning  space perturbations, such as gravity gradient effect, aerodynamic interference and solar radiation pressure torque. In this simulation, the measured terms can be specified as: $\hat{\bar m}=1.1\bar m$ may mean some space debris affixed to the subsatellite; $\bar L=1.02L$ and $\bar \Omega=0.98\Omega$ mean some measurement errors happening to the total length and the orbital angular velocity. In Fig~\ref{fig:6}, although the dynamics have already been  stable, the tension is switching in the high frequency. It is commonplace in the sliding mode control law design, called chattering, since the signum function is used in the input, which may cause unmodeled high frequency dynamics, and for reducing the chattering, we introduce the boundary layer technique into the proposed control laws with the thickness of the boundary layer specified as a positive scalar $\nu=2.5\times 10^{-3}$ in this simulation~\cite{1039802}.

The simulation results of the deployment dynamics subjected to uncertainties are exhibited from Fig~\ref{fig:7} to Fig~\ref{fig:10}. In Fig~\ref{fig:7}, although there exist parametric uncertainties compared with the dynamics in Figs~\ref{fig:3}-\ref{fig:6}, QST-SMC still performs the deployment with the shortest rise time and the minimum overshoot among the three methods, and the tether is deployed to the desired position at about 8 $rad$ by observing the dynamics of the in-plane and out-of-plane angles from Fig~\ref{fig:9}. In addition, the result in Fig~\ref{fig:10} shows that the tension falls in the reasonable region, and meanwhile the chattering of the input in the stable status is reduced by utilizing the boundary layer technique. The boundary layer works at about 6 rad as the error enters the specified boundary. Moreover, compared with the results in Fig~\ref{fig:6}, it is evident that the final value of the tension is consistent with the result of article~\cite{Ma201667}, since the discontinuous switching input is replaced by continuous input by the means of boundary layer. \textcolor{red}{Fig~\ref{fig:11} shows the in-plane and out-of-plane dimensionless thruster forces. From the results, we can find that all the forces are reasonable and limited in the boundary, and have less chattering by using the boundary layer technique. The thruster forces have the similar behaviors like the tension, and the chattering is almost eliminated in the final stage. Hence we can say that the boundary layer technique can indeed improve the inputs and repress the chattering.} The control precisions in Figs~\ref{fig:7}-\ref{fig:9} are almost same to Figs~\ref{fig:3}-\ref{fig:6}, which means the loss of precision caused by the boundary layer technique is limited. Moreover, the simulation results show that the good performance of proposed controller and the suitable boundary layer thickness guarantee that the system state error converges to the small vicinity of origin, which is consistent with the analyses in Theorem~\ref{thm:boundary}. Hence, the robustness of the proposed control laws can guarantee the stability of the deployment subjected to parametric uncertainties and external disturbances.
\includegraphics[width=0.6\textwidth]{paper4_fig7_20161025.eps}
\caption{Length dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:7}
\includegraphics[width=0.6\textwidth]{paper4_fig8_20161025.eps}
\caption{Length rate dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:8}
\includegraphics[width=0.6\textwidth]{paper4_fig9_20161025.eps}
\caption{Swing dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:9}
\includegraphics[width=0.6\textwidth]{paper4_fig10_20161025.eps}
\caption{Tension produced by utilizing boundary layer technique}
\label{fig:10}
\includegraphics[width=0.6\textwidth]{paper4_response_20161230.eps}
\caption{Dimensionless thruster force}
\label{fig:11}
\section{Conclusion}\label{sec:5}
In this paper, the proposed adaptive sliding mode control laws provide the deployment dynamics of the TSS with a simplified solution for the reaching phase, which ignores the nonlinear nature of the model, and the new dimensionless transformation subjected to the limited input guarantees the effectiveness of the auxiliary Lyapunov function with adaptive terms. The simulation results show that the proposed methods adjust the length of the tether and the swing dynamics as desired, although there exists the input limitation during the deployment. Also, the robustness to parametric uncertainties and external disturbances is verified reasonably. According to the results mentioned above, combining error dynamics design approach, adaptive approach and nonlinear sliding mode surface design technology may provide a further insight into regulation of the deployment dynamics of the TSS.
\section{Acknowledgment}
This work is partially supported by the National Natural Science Foundation of China (No. 61104112, 61673009, 61503101, 91438202).
\section{References}
\bibliography{paper4_ref}
\bibliographystyle{elsarticle-num}
\end{document}
